Bullet Data Preparation for LIME Diagnostics Paper

### 2020-09-25 code/02-data-preparation.Rmd
removed an extra space
trying to compare the data

@@

>>>>

## Comparing Training Datasets

Load the newest training data provided by Heike to determine if the same data:

```{r}
ccfs_withlands_raw = read.csv("../data/raw/CCFs_withlands.csv")
```

Adjust the raw data to match the structure of `bullet_train_raw`:

```{r}
ccfs_withlands_raw_expanded <- 
  ccfs_withlands_raw %>%
  rename("study1" = "study.x", "barrel1" = "barrel.x", "bullet1" = "bullet.x", "land1" = "land.x",
         "study2" = "study.y", "barrel2" = "barrel.y", "bullet2" = "bullet.y", "land2" = "land.y") %>%
  select(
    study1, barrel1, bullet1, land1, 
    study2, barrel2, bullet2, land2,
    all_of(rf_features), 
    match
  ) %>%
  mutate(across(c(study1, barrel1, bullet1, land1, study2, barrel2, bullet2, land2), factor))
```

```{r}
bullet_train_raw_expanded <- 
  bullet_train_raw %>%
  separate(land_id1, c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(land_id2, c("study2", "barrel2", "bullet2", "land2")) %>%
  mutate(across(where(is.character), factor)) %>%
  mutate(barrel1 = fct_relevel(barrel1, paste0("Br", 1:10)),
         barrel2 = fct_relevel(barrel2, paste0("Br", 1:10)))
```

Shows that `Hamby44` needs to be changed to `Hamby173` in `ccfs_withlands`:

```{r fig.width = 10, fig.height = 4, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Shows that lettered barrels need to be changed to `BrUnk` and `Br` added to other barrels in `ccfs_withlands`:

```{r fig.width = 10, fig.height = 4, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Shows that barrel letters from in `ccfs_withlands` need to be made the bullet and `B` added to all bullets:

```{r fig.width = 10, fig.height = 4, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Shows that `L` needs to be added to lands in `ccfs_withlands`:

```{r fig.width = 10, fig.height = 4, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

```{r}
# Determine the letters in barrels
letters <- 
  ccfs_withlands_raw_expanded %>%
  filter(!(barrel1 %in% 1:10)) %>% 
  pull(barrel1) %>%
  unique()

letters2 <- 
  ccfs_withlands_raw_expanded %>%
  filter(!(barrel2 %in% 1:10)) %>% 
  pull(barrel2) %>%
  unique()
sum(letters != letters2) > 0

ccfs_withlands_expanded <- ccfs_withlands_raw_expanded %>%
  mutate(
    study1 = fct_recode(study1, "Hamby173" = "Hamby44"),
    study2 = fct_recode(study2, "Hamby173" = "Hamby44"),
    bullet1 = ifelse(barrel1 %in% letters, as.character(barrel1), as.character(bullet1)),
    barrel1 = ifelse(barrel1 %in% letters, "Unk", barrel1),
    bullet2 = ifelse(barrel2 %in% letters, as.character(barrel2), as.character(bullet2)),
    barrel2 = ifelse(barrel2 %in% letters, "Unk", barrel2)
    ) %>%
  mutate(
    study1 = fct_relevel(study1, "Hamby173", "Hamby252"),
    study2 = fct_relevel(study2, "Hamby173", "Hamby252"),
    barrel1 = paste0("Br", barrel1),
    bullet1 = paste0("B", bullet1),
    land1 = paste0("L", land1),
    barrel2 = paste0("Br", barrel2),
    bullet2 = paste0("B", bullet2),
    land2 = paste0("L", land2)
  ) %>% 
  mutate(across(where(is.character), factor)) %>%
  mutate(barrel1 = fct_relevel(barrel1, paste0("Br", 1:10)),
         barrel2 = fct_relevel(barrel2, paste0("Br", 1:10)))
```

The labels agree now but note that :

```{r fig.width = 10, fig.height = 4, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

The labels agree now but we are starting to see some differences in the two datasets:

```{r fig.width = 10, fig.height = 4, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

The labels agree now but again, there are apparent differences between datasets:

```{r fig.width = 10, fig.height = 4, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_expanded %>% count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

The labels agree now:

```{r fig.width = 10, fig.height = 4, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_expanded %>% count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

```{r}
ccfs_withlands <- 
  ccfs_withlands_expanded %>%
  mutate(
    land_id1 = paste(study1, barrel1, bullet1, land1, sep = "-"),
    land_id2 = paste(study2, barrel2, bullet2, land2, sep = "-")
  ) %>%
  select(land_id1, land_id2, all_of(rf_features), match)
```

```{r fig.width = 12, fig.height = 6, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% 
    mutate(across(where(is.character), factor)) %>%
    filter(study1 == "Hamby173", study2 == "Hamby173") %>%
    filter(bullet1 %in% c("B1","B2"), bullet2 %in% c("B1", "B2")) %>% 
    count(study1, barrel1, bullet1, land1, study2, barrel2, bullet2, land2) %>%
    ggplot(aes(
      x = barrel1:land1, y = barrel2:land2, fill = n
    )) +
    geom_tile() +
    facet_grid(study2 + bullet2 ~ study1 + bullet1) +
    theme(aspect.ratio = 1,
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_expanded %>% 
    filter(study1 == "Hamby173", study2 == "Hamby173") %>%
    filter(bullet1 %in% c("B1","B2"), bullet2 %in% c("B1", "B2")) %>% 
    count(study1, barrel1, bullet1, land1, study2, barrel2, bullet2, land2) %>%
    ggplot(aes(
      x = barrel1:land1, y = barrel2:land2, fill = n
    )) +
    geom_tile() +
    facet_grid(study2 + bullet2 ~ study1 + bullet1) +
    theme(aspect.ratio = 1, 
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
    labs(title = "CCFs with lands data")
)
```

```{r}
bullet_train_raw_switch <-
  bullet_train_raw_expanded %>%
  mutate(across(where(is.factor), as.character)) %>%
  mutate(temp_barrel = ifelse(barrel1 == "Br10", barrel1, "NA")) %>%
  mutate(barrel1 = ifelse(temp_barrel == "Br10", barrel2, barrel1)) %>%
  mutate(barrel2 = ifelse(temp_barrel == "Br10", temp_barrel, barrel2)) %>%
  select(-temp_barrel) %>%
  mutate(across(where(is.character), factor)) %>%
  mutate(barrel1 = fct_relevel(barrel1, paste0("Br", 1:10)),
         barrel2 = fct_relevel(barrel2, paste0("Br", 1:10)))
```

```{r fig.width = 15, fig.height = 30, echo = FALSE}
plot_grid(
  bullet_train_raw_switch %>% 
    filter(study1 == "Hamby173", study2 == "Hamby173") %>%
    filter(bullet1 %in% c("B1","B2"), bullet2 %in% c("B1", "B2")) %>% 
    count(study1, barrel1, bullet1, study2, barrel2, bullet2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    facet_grid(study2 + bullet2 ~ study1 + bullet1) +
    theme(aspect.ratio = 1,
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_expanded %>% 
    filter(study1 == "Hamby173", study2 == "Hamby173") %>%
    filter(bullet1 %in% c("B1","B2"), bullet2 %in% c("B1", "B2")) %>% 
    count(study1, barrel1, bullet1, study2, barrel2, bullet2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    facet_grid(study2 + bullet2 ~ study1 + bullet1) +
    theme(aspect.ratio = 1, 
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
    labs(title = "CCFs with lands data"), 
bullet_train_raw_switch %>% 
    filter(study1 == "Hamby173", study2 == "Hamby252") %>%
    filter(bullet1 %in% c("B1","B2"), bullet2 %in% c("B1", "B2")) %>% 
    count(study1, barrel1, bullet1, study2, barrel2, bullet2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    facet_grid(study2 + bullet2 ~ study1 + bullet1) +
    theme(aspect.ratio = 1,
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_expanded %>% 
    filter(study1 == "Hamby173", study2 == "Hamby252") %>%
    filter(bullet1 %in% c("B1","B2"), bullet2 %in% c("B1", "B2")) %>% 
    count(study1, barrel1, bullet1, study2, barrel2, bullet2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    facet_grid(study2 + bullet2 ~ study1 + bullet1) +
    theme(aspect.ratio = 1, 
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
    labs(title = "CCFs with lands data"),
bullet_train_raw_switch %>% 
    filter(study1 == "Hamby252", study2 == "Hamby173") %>%
    filter(bullet1 %in% c("B1","B2"), bullet2 %in% c("B1", "B2")) %>% 
    count(study1, barrel1, bullet1, study2, barrel2, bullet2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    facet_grid(study2 + bullet2 ~ study1 + bullet1) +
    theme(aspect.ratio = 1,
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_expanded %>% 
    filter(study1 == "Hamby252", study2 == "Hamby173") %>%
    filter(bullet1 %in% c("B1","B2"), bullet2 %in% c("B1", "B2")) %>% 
    count(study1, barrel1, bullet1, study2, barrel2, bullet2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    facet_grid(study2 + bullet2 ~ study1 + bullet1) +
    theme(aspect.ratio = 1, 
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
    labs(title = "CCFs with lands data"),
  bullet_train_raw_switch %>% 
    filter(study1 == "Hamby252", study2 == "Hamby252") %>%
    filter(bullet1 %in% c("B1","B2"), bullet2 %in% c("B1", "B2")) %>% 
    count(study1, barrel1, bullet1, study2, barrel2, bullet2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    facet_grid(study2 + bullet2 ~ study1 + bullet1) +
    theme(aspect.ratio = 1,
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_expanded %>% 
    filter(study1 == "Hamby252", study2 == "Hamby252") %>%
    filter(bullet1 %in% c("B1","B2"), bullet2 %in% c("B1", "B2")) %>% 
    count(study1, barrel1, bullet1, study2, barrel2, bullet2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    facet_grid(study2 + bullet2 ~ study1 + bullet1) +
    theme(aspect.ratio = 1, 
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
    labs(title = "CCFs with lands data"), 
  ncol = 2
)
```

```{r}
bullet_train_cleaned <-
  bullet_train_raw_switch %>%
  mutate(
    land_id1 = paste(study1, barrel1, bullet1, land1, sep = "-"),
    land_id2 = paste(study2, barrel2, bullet2, land2, sep = "-")
  ) %>%
  select(land_id1, land_id2, rf_features)

ccfs_withlands_cleaned <- ccfs_withlands_expanded %>%
  mutate(
    land_id1 = paste(study1, barrel1, bullet1, land1, sep = "-"),
    land_id2 = paste(study2, barrel2, bullet2, land2, sep = "-")
  ) %>%
  select(land_id1, land_id2, rf_features)
```

```{r}
land_ids_old = bullet_train_cleaned %>% select(land_id1, land_id2)
land_ids_new = ccfs_withlands_cleaned %>% select(land_id1, land_id2)
dim(land_ids_old)
dim(land_ids_new)
```

```{r}
extra_leas_in_old <- 
  anti_join(x = land_ids_old, y = land_ids_new, by = c("land_id1", "land_id2")) %>%
  separate(land_id1, c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(land_id2, c("study2", "barrel2", "bullet2", "land2"))
dim(extra_leas_in_old)
```

```{r fig.height = 6, echo = FALSE}
extra_leas_in_old %>%
  count(study1, barrel1, study2, barrel2) %>%
  ggplot(aes(x = barrel1, y = barrel2, fill = n)) +
  geom_tile() +
  facet_grid(study2 ~ study1, scales = "free") + 
  theme(aspect.ratio = 1)
```

```{r}
extra_leas_in_new <- 
  anti_join(x = land_ids_new, y = land_ids_old, by = c("land_id1", "land_id2")) %>%
  separate(land_id1, c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(land_id2, c("study2", "barrel2", "bullet2", "land2"))
dim(extra_leas_in_new)
```

```{r fig.height = 6, echo = FALSE}
extra_leas_in_new %>%
  count(study1, barrel1, study2, barrel2) %>%
  ggplot(aes(x = barrel1, y = barrel2, fill = n)) +
  geom_tile() +
  facet_grid(study2 ~ study1, scales = "free") + 
  theme(aspect.ratio = 1)
```

```{r}
extra_leas_in_new <- 
  anti_join(x = land_ids_new, y = land_ids_old, by = c("land_id1", "land_id2")) %>%
  separate(land_id1, c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(land_id2, c("study2", "barrel2", "bullet2", "land2"))
dim(extra_leas_in_new)
```

@@

```{r}
bullet_train_eric_filter_v2 <-
  bullet_train_raw %>%
  mutate(bbl1 = str_remove(land_id1, pattern = "Hamby252-|Hamby173-"),
         bbl2 = str_remove(land_id2, pattern = "Hamby252-|Hamby173-")) %>%
  filter(!(bbl1 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4") | 
           bbl2 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4"))) %>%
  select(-bbl1, -bbl2) %>%
  separate(col = "land_id1",
           into = c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(col = "land_id2",
           into = c("study2", "barrel2", "bullet2", "land2")) %>%
  filter(study1 != "Hamby173" | bullet1 != "BE") %>%
  filter(study2 != "Hamby173" | bullet2 != "BE")

dim(bullet_train_eric_filter_v2)
length(bulletxtrctr::rtrees$predicted) == dim(bullet_train_eric_filter_v2)[1]
```
